---
title: "Hurricane Mapping from FEMA"
author: "Helen"
date: "11-10-2020"
output:
  revealjs::revealjs_presentation:
    theme: solarized
    highlight: pygments
    css: leaflet-reveal.css
    center: false
    transition: slide
---


```{r setup, include=FALSE}
library(knitr)
library(lubridate)
library(stringr)
library(dplyr)
library(magrittr)
library(tigris)
library(rgdal)
library(geojsonio)
library(leaflet)
library(kableExtra)
```

## explanations of data
<font size ="5"> 
This assignment use the  Public Assistance Funded Projects Details dataset from FEMA.
</font>

<font size ="4"> 
The data were <a href="https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1">stored online</a > and then downloaded as a CSV file.</p >
</font>

## display the data 
The table below shows the data heading used for mapping.
```{r,echo=FALSE}
public.new<-read.csv("public.new2.csv")
#summarize with each fips, delete the dubricate lines of states and counties
display<- public.new %>% group_by(fips) %>%summarise(county=unique(county),state=unique(state), incidentType=unique(incidentType),n=n(), sumObligated=sum(totalObligated)) %>% rename(GEO_ID = fips)
#round the amount with thousands
display$sumObligated<- display$sumObligated/1000
kable(head(display),"html") %>%
  kable_styling(font_size = 15)
```
<font size ="4"> 
<p align="left">
-GEO_ID: fips code of the county

-state: state name

-county: county name

-n:numbers of the funded hurricane projects between 2009-2018

-sumObligated: Sum of the federal share of the Public Assistance grant eligible project amount in dollars, plus grantee (State) and sub-grantee (applicant) administrative costs between 2009-2018.

-incidentType: Type of incident:"Hurricane","Severe Storm(s)","Coastal Storm"
</p>
</font>

## mappings
<h4> The map shows the sum of Obligated amount for the hurricane disasters between 2009-2018.</h4>
```{r, echo=FALSE}
#get the layer of US map from the ".json" file as xy
#The data comes from JSON.
#We’ll use the geojsonio package to load the data from javascrpt into sp objects, which will let us easily manipulate the geographic features, and their properties, in R.
xy <- geojsonio::geojson_read("gz_2010_us_050_00_5m.json", what = "sp")
#names(xy)
#get the string between 10-14 digit of GEO_ID to match the 5-digit string of "fips" 
xy$GEO_ID %<>% substr(start = 10, stop = 14)

# join the US map data with funding data by fips/GEO-ID
leafmap <- geo_join(xy, display, by = "GEO_ID", how = "inner")

#pal function:maps data values "sumObligated" to colors according to a given palette
#pal: colors for the fund amount of hurricanes
pal <- leaflet::colorNumeric(palette = "Blues", domain = leafmap$sumObligated)
#paln: colors for the number of hurricanes
paln<- leaflet::colorNumeric(palette = "Reds", domain = leafmap$n)


#add labels
labels <- sprintf(
  "%s<br/>%s<br/>%s",
  leafmap$state,leafmap$county, leafmap$sumObligated) 
#add popups
popup<- paste0("<strong>state:</strong>", leafmap$state,"<br>",
               "<strong>county:</strong>", leafmap$county,"<br>",
               "<strong>Sum of Obligated Amount:</strong>",round(leafmap$sumObligated,2),"thousand")
popupn<- paste0("<strong>state:</strong>", leafmap$state,"<br>",
                "<strong>county:</strong>", leafmap$county,"<br>",
                "<strong>Number of funded hurricane projects:</strong>",leafmap$n)

#mapping by leaflet
leaflet() %>%
  #titles
  addTiles() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addProviderTiles(providers$Stamen.TonerLines,
                   options = providerTileOptions(opacity = 0.75)) %>%
  
  setView(-89.275673, 37.098, zoom = 4) %>%
  addPolygons(data = leafmap, 
              fillColor = ~pal(sumObligated), 
              color = "#BDBDC3",
              fillOpacity = 1, 
              weight = 1, 
              smoothFactor = 0.2,
              
              popup= popup,
              popupOptions = popupOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")
              
              ) %>%
  addLegend(pal = pal, 
            values = leafmap$sumObligated, 
            position = "bottomright", 
            title = paste0("Sum obligated amount(in thousands)<br>2009-2018")) 
```

------
<h4> The map shows the numbers of hurricanes funded projects between 2009-2018</h4> 

```{r,echo=FALSE}
leaflet() %>%
  #titles
  addTiles() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addProviderTiles(providers$Stamen.TonerLines,
                   options = providerTileOptions(opacity = 0.75)) %>%
  
  setView(-89.275673, 37.098, zoom = 4) %>%
  addPolygons(data = leafmap, 
              fillColor = ~paln(n), 
              color = "#BDBDC3",
              fillOpacity = 1, 
              weight = 1, 
              smoothFactor = 0.2,
              
              popup= popupn,
              popupOptions = popupOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")
              
  ) %>%
  addLegend(pal = paln, 
            values = leafmap$n, 
            position = "bottomright", 
            title = paste0("Numbers of funded hurricane projects<br>2009-2018"))

```
